Accelerating Neural Network Training using Arbitrary Precision Approximating Matrix Multiplication Algorithms

### 2021-03-07 latex/MPP21/paper.tex
incorporated C_11 example from Bini's alg

For a concrete example, we reproduce the rule for the algorithm developed by Bini et al.~\cite{BCRL79} for the $\dims{3,2,2}$ case (multiplying a $3\times 2$ matrix $A$ by a $2\times 2$ matrix $B$), where we use the following notation for input and output matrices:
$$
\begin{bmatrix} A_{11} & A_{12} \\ A_{21} & A_{22} \\ A_{31} & A_{32}  \end{bmatrix} \cdot
\begin{bmatrix} B_{11} & B_{12} \\ B_{21} & B_{22} \end{bmatrix}=
\begin{bmatrix} C_{11} & C_{12} \\ C_{21} & C_{22} \\ C_{31} & C_{32} \end{bmatrix} .$$
Bini's rule is given by
\begin{align*}
M_1 =& (A_{11} + A_{22}) \cdot (\lambda B_{11} + B_{22}) \\
M_2 =& A_{22}\cdot  (-B_{21} - B_{22}) \\
M_3 =& A_{11}\cdot B_{22} \\
M_4 =& (\lambda A_{12} + A_{22})\cdot (-\lambda B_{11} + B_{21}) \\
M_5 =& (A_{11} + \lambda A_{12}) \cdot (\lambda B_{12} + B_{22}) \\
M_6 =& (A_{21} + A_{32}) \cdot (B_{11} + \lambda B_{22}) \\
M_7 =& A_{21} \cdot (-B_{11} - B_{12}) \\
M_8 =& A_{32} \cdot B_{11}\\
M_9 =& (A_{21} + \lambda A_{31}) \cdot (B_{12} - \lambda B_{22})\\
M_{10} =& (\lambda A_{31} + A_{32}) \cdot (B_{12} - \lambda B_{22})\\
\end{align*}
\begin{align*}
\hat C_{11} =& \lambda^{-1} (M_1 + M_2 - M_3 + M_4) \\
\hat C_{12} =& \lambda^{-1} (-M_3 + M_5) \\
\hat C_{21} =& M_4 + M_6 - M_{10}\\
\hat C_{22} =& M_1 - M_5 + M_9\\
\hat C_{31} =& \lambda^{-1} (-M_8 + M_{10}) \\
\hat C_{32} =& \lambda^{-1} (M_6 + M_7 - M_8 + M_9).\\
\end{align*}

\GB{give explicit expression for $\hat{C}_{11}$ and reference \cref{eq:APAapprox}}
We highlight several properties of the rule which are common across all APA algorithms we consider.
First, the rule requires fewer multiplications than the classical one (rank 10 instead of 12 for $\dims{3,2,2}$ here).
Next, each multiplication is between a linear combination of entries of $A$ and a linear combination of entries of $B$, each output entry is computed as a linear combination of the outputs of the multiplications, and each coefficient in the linear combinations is a (Laurent) polynomial in $\lambda$.
The coefficients include both positive and negative powers of $\lambda$, which explains why small values of $\lambda$ can lead to significant roundoff error.
For Bini's algorithm all coefficients are monomial with degree between $-1$ and $1$.

>>>>

For a concrete example, we reproduce the rule for the algorithm developed by Bini et al.~\cite{BCRL79} for the $\dims{3,2,2}$ case (multiplying a $3\times 2$ matrix $A$ by a $2\times 2$ matrix $B$), where we use the following notation for input and output matrices:
$$
\begin{bmatrix} A_{11} & A_{12} \\ A_{21} & A_{22} \\ A_{31} & A_{32}  \end{bmatrix} \cdot
\begin{bmatrix} B_{11} & B_{12} \\ B_{21} & B_{22} \end{bmatrix}=
\begin{bmatrix} C_{11} & C_{12} \\ C_{21} & C_{22} \\ C_{31} & C_{32} \end{bmatrix} .$$
Bini's rule is given by
\begin{align*}
M_1 =& (A_{11} + A_{22}) \cdot (\lambda B_{11} + B_{22}) \\
M_2 =& A_{22}\cdot  (-B_{21} - B_{22}) \\
M_3 =& A_{11}\cdot B_{22} \\
M_4 =& (\lambda A_{12} + A_{22})\cdot (-\lambda B_{11} + B_{21}) \\
M_5 =& (A_{11} + \lambda A_{12}) \cdot (\lambda B_{12} + B_{22}) \\
M_6 =& (A_{21} + A_{32}) \cdot (B_{11} + \lambda B_{22}) \\
M_7 =& A_{21} \cdot (-B_{11} - B_{12}) \\
M_8 =& A_{32} \cdot B_{11}\\
M_9 =& (A_{21} + \lambda A_{31}) \cdot (B_{12} - \lambda B_{22})\\
M_{10} =& (\lambda A_{31} + A_{32}) \cdot (B_{12} - \lambda B_{22})\\
\end{align*}
\begin{align*}
\hat C_{11} =& \lambda^{-1} (M_1 + M_2 - M_3 + M_4) \\
\hat C_{12} =& \lambda^{-1} (-M_3 + M_5) \\
\hat C_{21} =& M_4 + M_6 - M_{10}\\
\hat C_{22} =& M_1 - M_5 + M_9\\
\hat C_{31} =& \lambda^{-1} (-M_8 + M_{10}) \\
\hat C_{32} =& \lambda^{-1} (M_6 + M_7 - M_8 + M_9).\\
\end{align*}
For example, this rule computes the first entry of the output matrix as $\hat C_{11} = A_{11}  B_{11} + A_{12}  B_{21} - \lambda A_{12}  B_{11}$.
Thus, in this case, the first entry of the error matrix $E$ in \cref{eq:APAapprox} is $E_{11} = A_{12}  B_{11}$.

We highlight several properties of the rule which are common across all APA algorithms we consider.
First, the rule requires fewer multiplications than the classical one (rank 10 instead of 12 for $\dims{3,2,2}$ here).
Next, each multiplication is between a linear combination of entries of $A$ and a linear combination of entries of $B$, each output entry is computed as a linear combination of the outputs of the multiplications, and each coefficient in the linear combinations is a (Laurent) polynomial in $\lambda$.
The coefficients include both positive and negative powers of $\lambda$, which explains why small values of $\lambda$ can lead to significant roundoff error.
For Bini's algorithm all coefficients are monomial with degree between $-1$ and $1$.
